\chapter*{Acknowledgements}

\newcommand{\gf}{fianc\'{e}e}

First and foremost, I would like to thank my supervisor Alan Fekete for his expertise and knowledge in the research discipline, providing me with plenty of guidance, feedback and always making time to meet with me. His critical eye and invaluable feedback right up until this thesis deadline have been crucial to the success of my thesis and the betterment to my understanding of the research community as a whole. Without his support, I would not have been able to achieve the level and quality of research I managed to attain. 

I would also like to thank Michael Cahill from MongoDB for giving me the initial idea for the project and assisting me throughout the year with his indepth knowledge of MongoDB and how best to approach tackling this research. I would like to thank the Australian Research Council for providing the School of Information Technologies and MongoDB with a Linkage grant.

A thank you goes out to Julia Wong for persistently reaching out and making sure this thesis is on track. Your continuous proofreading, feedback and suggestions were an immense help throughout the year. Thank you to Alison Wong, Nicky Ringland and Shane Arora for proofreading my thesis and pointing out the improvements I could make. Your keen eyes were a huge help in getting this thesis to the finish line. 

Super special thanks go to my \gf \ Deanna Arora for her continual support in the form sweets, support and general motivation throughout the year. Thank you for being there for me through the thick and thin of writing code, running experiments, writing this thesis and proofreading it until its completion.